Long range correlations and their event scale dependence in high multiplicity pp collisions at $\sqrt{s} = 13$~TeV

### 2020-06-22 sec_results.tex
sec5.2

Figure~\ref{fig:PlotCorrMBHMT} shows two-particle correlation functions as functions of the relative pseudorapidity ($\Delta \eta$) and azimuthal angle ($\Delta \varphi$) differences between the trigger and associated particles. The functions are measured with Eq.~\ref{eq:corrfunction} for $1 < \pttrig~\mathrm{and}~\ptassoc < 2$ GeV/$c$ in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV}. The left figure is for the 0--0.1\% multiplicity class. The right one is from the minimum bias events (0--100\% multiplicity class). The ridge structure is clearly observed in the high multiplicity class compared to the minimum-bias events while no observation of the ridge in the minimum bias events. The away side is dominantly constructed by correlation of back to back jet.

@@

 Projected $\Delta\varphi$ distributions of the two-particle correlation functions in $1.6<|\Delta\eta|<1.8$ are shown in Fig.~\ref{fig:PlotDeltaPhi} for various $\it{p}_{\rm{T}}$ intervals in the minimum bias class(0--100\%) and the high multiplicity class (0--0.1\%). The near-side($\Delta\varphi\sim 0$) peak in the minimum bias class is rarely seen in the various transverse momentum range. The near-side peak is highest in the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gradually decreases with increasing $\it{p}_{\rm{T}}$ in the high multiplicity class. The measurements in the high multiplicity class are compared with the CMS results. The near-side peak measured with CMS shows agreement with ALICE measurement, while showing higher yield in the away-side($\Delta\varphi\sim\pi$) peak. The model predictions are represented for the measurements in the high multiplicity class. The $\pythiashoving$ nicely estimates the near-side peak and overestimates the away-side yield fir the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$. The $\pythiashoving$ does not reproduce the near-side peak for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range. $\pythiam$ does not show any peaks in the near-side, which is expected. While not reproducing the near-side peak, $\pythiam$ estimates the away-side peak, which mainly come from the jet--balancing jet correlations.

@@

The spectra of the ridge yield are shown in Fig.~\ref{fig:PlotYSpect} in the high multiplicity class and compared with the CMS results~\cite{Khachatryan:2015lva}. The estimator of particle multiplicity of ALICE is done with forward subsystem(V0), whereas that of CMS is done by mid-rapidity particles meeting with the condition of $|\eta|<$2.4 and $\pt>0.4$ GeV/$c$. Dedicated comparison is conducted and the difference of particle multiplicity is estimated to be about 20\%. Taking into account the difference in acceptance of charged tracks and comparable definition of multiplicity, the measurements are can be considered comparable with each other, which insists that measurements of azimuthal correlations can be achieved with ALICE. The spectrum is also compared with model predictions. $\pythiam$ and $\pythiashoving$ are used for the prediction. $\pythiam$ is shown with a orange line and predicts no ridge effect in pp collisions, which is expected because the model does not implement any constraint to describe the ridge effect. $\pythiashoving$ shown with a blue line describes the effect better than $\pythiam$, however, its yield decreases more rapidly than the data as $\pttrigassoc$ increases.

@@

\subsection{Event scale dependent ridge yield}

To further observe the behavior of the ridge yield with respect to the hard process, the correlation function is measured with the minimum $\ptlead$ (left) or $\ptjet$ (right) thresholds for $1<\pttrigassoc<2$ GeV/$c$, where the measured ridge yield is maximum in Fig.~\ref{fig:PlotCorrHMTSel}. The ridge structures are still seen with the event scale selection. With the minimum $\ptjet$ selection, the correlations function has a bump on the $\Delta\eta = 0$ and $\Delta\varphi = \pi$, which occurs due to the limited jet acceptance.

@@

The one-dimensional $\Delta\varphi$ distribution with the minimum $\ptlead$ (left) or $\ptjet$ is shown in Fig.~\ref{fig:PlotDeltaPhiESE}. The near-side yield increases as the event scale increases by requiring higher $\ptlead$ (left) or $\ptjet$. The model comparisons are are shown for $\pythiashoving$ with a blue and $\pythiam$ with a red line. The $\pythiashoving$ describes ridge yield with the selection. $\pythiam$ predicts non ridge yield with the selection, which also tell there is little ignore-able jet contamination. Both models well describe the away-side yield with the $\ptjet$ selection and overestimates away-side yield with the $\ptlead$ selection.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig5_PlotDeltaPhiESE.pdf}
	\caption{ One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ with the minimum $\ptlead$(lower) or $\ptjet$(upper) selection for $1<\pttrigassoc<2$ GeV/\it{c}\rm{} in the 0--0.1\% multiplicity class. The filled circles show measurement with ALICE. Model predictions are compared with measurement with a blue line for $\pythiashoving$  and a orange line for $\pythiam$.}
	\label{fig:PlotDeltaPhiESE}
\end{figure}

The spectra of the ridge yield as function of the minimum $\it{p}_{\rm{T, Lead}}$(left) or $\it{p}_{\rm{T, Jet}}$(right) are shown in Fig.~\ref{fig:RidgeYield_ESE}. The spectra are compared with model predictions. 
The increases of ridge yield with the selections are observed and the model with PYTHIA8 + String Shoving underestimates the ridge yield. with the $\it{p}_{\rm{T, Jet}}$ selection. the model prediction with PYTHIA8 default still assures ignorable jet contamination.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig6_RidgeYieldESE.pdf}
	\caption{The ridge yield spectra with respect to the leading particle and jet selections. The filled circles show measurement with ALICE. Model predictions are compared with measurement as blue line for PYTHIA8 + String Shoving  and apricot line for PYTHIA8 default.}
	\label{fig:RidgeYield_ESE}
\end{figure}

>>>>

Figure~\ref{fig:PlotCorrMBHMT} shows two-particle correlation functions as functions of the relative pseudorapidity ($\Delta \eta$) and azimuthal angle ($\Delta \varphi$) differences between the trigger and associated particles. The functions are measured with Eq.~\ref{eq:corrfunction} for $1 < \pttrig~\mathrm (\ptassoc) < 2$ GeV/$c$ in pp collisions at $\sqrt{\it{s}} = $\unit{13} {\rm{}TeV}. The left panel is for the 0--0.1\% multiplicity class. The right one is from the minimum bias events (0--100\% multiplicity class). The ridge structure is clearly observed in the high multiplicity class compared to the minimum-bias events while it is not seen in the minimum bias events. The away side is dominant from the correlation of back to back jet.

@@

 Projected $\Delta\varphi$ distributions of the two-particle correlation functions in $1.6<|\Delta\eta|<1.8$ are shown in Fig.~\ref{fig:PlotDeltaPhi} for various $\it{p}_{\rm{T}}$ intervals in the minimum bias class(0--100\%) and the high multiplicity class (0--0.1\%). The near-side($\Delta\varphi\sim 0$) peak in 0--0.1\% multiplicity class is clearly observed in all the transverse momentum ranges studied while there is no hint of signal in the minimum bias class. The near-side peak is highest in the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gradually decreases with increasing $\it{p}_{\rm{T}}$ in the high multiplicity class. The measurements in the high multiplicity class are compared with the results measured by the CMS collaboration~\cite{Khachatryan:2015lva}. The near-side peaks in all transverse momentum ranges are comparable each other within the uncertainties. The larger away yields observed in \cite{Khachatryan:2015lva} can be attributed to the difference in $\eta$ acceptances. The model predictions are presented for the measurements in the high multiplicity class with the same acceptances used in this article both for $\eta$ and event multiplicity. The $\pythiashoving$ gives good estimates of the near-side peak and slightly overestimates the away-side yield for the interval of $1<\it{p}_{\rm{T}}<2$ GeV/$c$. However the $\pythiashoving$ underestimates the near-side peak for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range. $\pythiam$ does not show any peaks in the near-side as expected and slightly underestimates the away-side peak for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ and gives good estimates for the $\it{p}_{\rm{T}}>2$ GeV/$c$ range, which mainly come from the back to back jet correlations.

@@

The ridge yields as a function of  the transverse momentum are shown in Fig.~\ref{fig:PlotYSpect} in the high multiplicity class and compared to \cite{Khachatryan:2015lva}. The particle multiplicity in this analysis is estimated with the forward subsystem(V0), whereas that in \cite{Khachatryan:2015lva}  used the mid-rapidity particles in $|\eta|<$2.4 and $\pt>0.4$ GeV/$c$. By interpolating the $\eta$ and $\pt{}$ distributions based on PYTHIA, it is found that the multiplicity in \cite{Khachatryan:2015lva} is about 20\% higher than 0--0.1\% in this analysis, corresponding 0--0.03\%. Taking into account the differences in acceptance of charged tracks and estimated multiplicity, the measurements are comparable with each other. The spectrum is also compared with model calculations. $\pythiam$ gives zero yields since it does not contain the ridge effect. $\pythiashoving$ describes the yield qualitatively however its yield decreases more rapidly than the data as $\pttrigassoc$ increases.

@@

\subsection{Event scale dependent ridge yield}
The event scale dependent ridge yields are measured with the minimum $\ptlead=9.0$ (left) or $\ptjet=10.0$~ GeV/$c$ (right) thresholds for $1< \pttrig~\mathrm (\ptassoc) <2$ GeV/$c$, where the measured ridge yield is maximum without event scale selection. In Fig.~\ref{fig:PlotCorrHMTSel}, The ridge structure still persist even with both selections. With the minimum $\ptjet$ selection, the correlations function has double peaks on the $\Delta\eta$ at $\Delta\varphi = \pi$ due to the requirement of the limited jet acceptance.

@@

Projected $\Delta\varphi$ distributions of the correlation functions in $1.6<|\Delta\eta|<1.8$ with the minimum $\ptlead$(lower) and $\ptjet$(upper) are shown in Fig.~\ref{fig:PlotDeltaPhiESE}. The near-side yield increases as requiring higher $\ptlead$ or $\ptjet$ from the yield in unbiased events, which is shown in the upper left panel. The Model calculations by $\pythiashoving$ and $\pythiam$ are shown with a blue and a orange line, respectively. The $\pythiashoving$ gives qualitative agreement of the near-side peak with the event scale selections. On the other hand, $\pythiam$ does not show the near-side peak with the event scale selections. Both models describe the away-side yield with the $\ptjet$ selection and overestimate away-side yield with the $\ptlead$ selection.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig5_PlotDeltaPhiESE.pdf}
	\caption{ One-dimensional $\Delta\varphi$ distribution in the large $\Delta\eta$ with the minimum $\ptlead$(lower) or $\ptjet$(upper) selection for $1< \pttrig~\mathrm (\ptassoc) <2$ GeV/$c$ in the 0--0.1\% multiplicity class. The filled circles show measurement with ALICE. Model predictions are compared with measurement with a blue line for $\pythiashoving$  and a orange line for $\pythiam$.}
	\label{fig:PlotDeltaPhiESE}
\end{figure}

The ridge yields as functions of the minimum $\ptlead$(left) and $\ptjet$(right) selections are shown in Fig.~\ref{fig:RidgeYield_ESE}. The ridge yield increases with the increasing selection requirement.$\pythiashoving$ underestimates the ridge yield with the increasing event scale selection. $\pythiam$ does not produce the ridge yield, estimating zero yield, which imply the negligible jet contamination in the measurements.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.99\linewidth]{./figures/Fig6_RidgeYieldESE.pdf}
	\caption{The ridge yield spectra with respect to the leading particle and jet selections. The filled circles show measurement with ALICE. Model predictions are compared with measurement as blue line for $\pythiashoving$ and orange line for $\pythiam$.}
	\label{fig:RidgeYield_ESE}
\end{figure}
